---
title: "10_FUMA_val_genes_tileplots"
author: "Heather E. Wheeler"
date: '`r Sys.time()`'
output: html_document
---

```{r}
library(dplyr)
library(data.table)
library(ggplot2)
library(ggrepel)
library(gridExtra)
"%&%" = function(a,b) paste(a,b,sep="")
date <- Sys.Date()
```


```{r,fig.width=10,fig.height=10, message=FALSE}
my.dir <-  "/Volumes/im-lab/nas40t2/hwheeler/trans-px/NCBI_Gene_Summaries/paper_figures_2018-10-26/"
mtres <- fread(my.dir %&% 'TableS2_MultiXcan_results_2018-10-29.txt')
#rm version number from ensembl id

mtres <- mutate(mtres,neglog=ifelse(FHS_Pval<1e-50,50,-log10(FHS_Pval)),obsChr=ifelse(obsChr=="X",23,obsChr)) %>% mutate(obsChr=as.integer(obsChr))
sorted <- arrange(mtres,predChr,predS1,obsChr,obsS1)
#arrange by obs (target) for hvec
resobssort <- dplyr::arrange(mtres,obsChr,obsS1,predChr,predS1)
```

- Try with chromosomal coordinates instead of gene order
```{r,fig.width=7,fig.height=7, message=FALSE}
#get chromosomal coordinates
d <- sorted
predticks = NULL
predlastbase=0
d$predpos=NA
for (i in unique(d$predChr)) {
  if (i==1) {
  	d[d$predChr==i, ]$predpos=d[d$predChr==i, ]$predS1
  } else {
  	predlastbase=predlastbase+tail(subset(d,predChr==i-1)$predS1, 1)
  	d[d$predChr==i, ]$predpos=d[d$predChr==i, ]$predS1+predlastbase
  }
#  predticks=c(predticks, d[d$predChr==i, ]$predpos[floor(length(d[d$predChr==i, ]$predpos)/2)+1])
  predticks = c(predticks, predlastbase)
}
predticks = c(predticks, max(d$predpos))
#get chromosomal coordinates
e <- resobssort
obsticks = NULL
obslastbase=0
e$obspos=NA
for (i in unique(e$obsChr)) {
  if (i==1) {
  	e[e$obsChr==i, ]$obspos=e[e$obsChr==i, ]$obsS1
  } else {
  	obslastbase=obslastbase+tail(subset(e,obsChr==i-1)$obsS1, 1)
  	e[e$obsChr==i, ]$obspos=e[e$obsChr==i, ]$obsS1+obslastbase
  }
  obsticks = c(obsticks,obslastbase)
}
obsticks = c(obsticks,max(e$obspos))
esubset <- dplyr::select(e,obsgene,predgene,obspos)
d <- left_join(d,esubset,by=c("obsgene","predgene"))

#get position for chr labels
predchrticks <- c()
for(i in 2:length(predticks)){
  midpt <- mean(c(predticks[i-1],predticks[i]))
  predchrticks <- c(predchrticks,midpt)
}
obschrticks <- c()
for(i in 2:length(obsticks)){
  midpt <- mean(c(obsticks[i-1],obsticks[i]))
  obschrticks <- c(obschrticks,midpt)
}

#only include validated genes (P<0.05 in DGN)
d <- filter(d,DGN_Pval<0.05)
```

# Trans-acting pathway enrichment
## color by GO nucleic acid binding transcription factor activity
- only validated genes are shown
```{r,fig.width=5,fig.height=5, message=FALSE}
txpath <- c("RFX5", "ZNF438", "ZNF641", "ZSCAN10", "ZNF200", "ZNF263", "ZNF174", "TCF25", "GATAD2A", "ZFP14", "ZFP82", "ZNF260", "ZNF382", "ZNF790", "ZNF345", "ZNF829", "ZNF568", "ZNF585A", "ZNF570", "ZNF793", "ZNF540", "ZFP30", "ZNF607", "ZNF573", "ZNF649", "ZNF613", "ZNF132", "TEF", "PHF5A", "ZBTB11", "REST", "CITED2", "PLAGL1", "ZNF12")
sorted2 <- mutate(d,txpath=ifelse(d$predname %in% txpath,'yes','no'))
ggplot(sorted2,aes(x=predpos, y=obspos, size=neglog, color=txpath)) + 
  geom_point(shape=21)  +  
  scale_size_continuous(guide=FALSE) +
  scale_colour_manual(values=c("dark gray","blue"),guide=FALSE) +
  geom_hline(yintercept=obsticks,size=0.2) + 
  geom_vline(xintercept = predticks,size=0.2) + ggtitle('A') +
  xlab('trans-acting gene') + ylab('target gene') + 
  coord_cartesian(xlim=c(predticks[1],predticks[23]),ylim=c(obsticks[1],obsticks[24]),expand=FALSE) +
  geom_text_repel(data=dplyr::filter(sorted2,txpath=='yes'),aes(label=paste(predname,":",obsname,sep=""))) + theme_classic(12) +
  scale_x_continuous(breaks=predchrticks,labels= c(1:16,"",18,"",20,"",22)) +
  scale_y_continuous(breaks=obschrticks,labels=c(1:16,"",18,"",20,"",22,"X"))

sfig1a <- ggplot(sorted2,aes(x=predpos, y=obspos, size=neglog, color=txpath)) + 
  geom_point(shape=21)  +  
  scale_size_continuous(guide=FALSE) +
  scale_colour_manual(values=c("dark gray","blue"),guide=FALSE) +
  geom_hline(yintercept=obsticks,size=0.2) + 
  geom_vline(xintercept = predticks,size=0.2) + ggtitle('GO nucleic acid binding transcription factor activity') +
  xlab('trans-acting gene') + ylab('target gene') + 
  coord_cartesian(xlim=c(predticks[1],predticks[23]),ylim=c(obsticks[1],obsticks[24]),expand=FALSE) +
  geom_text_repel(data=dplyr::filter(sorted2,txpath=='yes'),aes(label=predname)) + theme_classic(10) +
  scale_x_continuous(breaks=predchrticks,labels= c(1:16,"",18,"",20,"",22)) +
  scale_y_continuous(breaks=obschrticks,labels=c(1:16,"",18,"",20,"",22,"X"))
```



#color by Reactome generic transcription pathway
```{r,fig.width=5,fig.height=5, message=FALSE}
txpath <- c("ZNF641", "ZNF200", "ZNF263", "NCOR1", "MED24", "ZNF566", "ZNF382", "ZNF585A", "ZNF569", "ZNF570", "ZNF540", "ZNF607", "ZNF649", "ZNF613", "ZNF12")
sorted2 <- mutate(d,txpath=ifelse(d$predname %in% txpath,'yes','no'))
ggplot(sorted2,aes(x=predpos, y=obspos, size=neglog, color=txpath)) + 
  geom_point(shape=21)  +  
  scale_size_continuous(guide=FALSE) +
  scale_colour_manual(values=c("dark gray","blue"),guide=FALSE) +
  geom_hline(yintercept=obsticks,size=0.2) + 
  geom_vline(xintercept = predticks,size=0.2) + ggtitle('Reactome generic transcription pathway') +
  xlab('trans-acting gene') + ylab('target gene') + 
  coord_cartesian(xlim=c(predticks[1],predticks[23]),ylim=c(obsticks[1],obsticks[24]),expand=FALSE) +
  geom_text_repel(data=dplyr::filter(sorted2,txpath=='yes'),aes(label=paste(predname,":",obsname,sep=""))) + theme_classic(12) +
  scale_x_continuous(breaks=predchrticks,labels= c(1:16,"",18,"",20,"",22)) +
  scale_y_continuous(breaks=obschrticks,labels=c(1:16,"",18,"",20,"",22,"X"))

sfig1b <- ggplot(sorted2,aes(x=predpos, y=obspos, size=neglog, color=txpath)) + 
  geom_point(shape=21)  +  
  scale_size_continuous(guide=FALSE) +
  scale_colour_manual(values=c("dark gray","blue"),guide=FALSE) +
  geom_hline(yintercept=obsticks,size=0.2) + 
  geom_vline(xintercept = predticks,size=0.2) + ggtitle('Reactome generic transcription pathway') +
  xlab('trans-acting gene') + ylab('target gene') + 
  coord_cartesian(xlim=c(predticks[1],predticks[23]),ylim=c(obsticks[1],obsticks[24]),expand=FALSE) +
  geom_text_repel(data=dplyr::filter(sorted2,txpath=='yes'),aes(label=predname)) + theme_classic(12) +
  scale_x_continuous(breaks=predchrticks,labels= c(1:16,"",18,"",20,"",22)) +
  scale_y_continuous(breaks=obschrticks,labels=c(1:16,"",18,"",20,"",22,"X"))
```

# Target Gene Pathway Enrichment

## Color by TF targets (MsigDB c3)	WGGAATGY_TEF1_Q6
```{r,fig.width=5,fig.height=5, message=FALSE}
txpath <- c("NID1", "VCL", "PRRG4", "TSPAN9", "PTMS", "DACT1", "SPTB", "ROGDI", "GHRL", "PROK2", "SPARC", "SH3BGRL2", "FBXL13", "TSPAN33")
sorted2 <- mutate(d,txpath=ifelse(d$obsname %in% txpath,'yes','no'))
sfig2a <- ggplot(sorted2,aes(x=predpos, y=obspos, size=neglog, color=txpath)) + 
  geom_point(shape=21)  +  
  scale_size_continuous(guide=FALSE) +
  scale_colour_manual(values=c("dark gray","red3"),guide=FALSE) +
  geom_hline(yintercept=obsticks,size=0.2) + 
  geom_vline(xintercept = predticks,size=0.2) + ggtitle('TF binding site: WGGAATGY_TEF1_Q6') +
  xlab('trans-acting gene') + ylab('target gene') + 
  coord_cartesian(xlim=c(predticks[1],predticks[23]),ylim=c(obsticks[1],obsticks[24]),expand=FALSE) +
  geom_text_repel(data=dplyr::filter(sorted2,txpath=='yes'),aes(label=paste(predname,":",obsname,sep=""))) + theme_classic(12) +
  scale_x_continuous(breaks=predchrticks,labels= c(1:16,"",18,"",20,"",22)) +
  scale_y_continuous(breaks=obschrticks,labels=c(1:16,"",18,"",20,"",22,"X"))
```


## Color by TF targets (MsigDB c3)	PAX8_B
```{r,fig.width=5,fig.height=5, message=FALSE}
txpath <- c("TSPAN2", "NRGN", "PPM1A", "ABLIM3", "CALD1", "BNC2")
sorted2 <- mutate(d,txpath=ifelse(d$obsname %in% txpath,'yes','no'))
sfig2c <- ggplot(sorted2,aes(x=predpos, y=obspos, size=neglog, color=txpath)) + 
  geom_point(shape=21)  +  
  scale_size_continuous(guide=FALSE) +
  scale_colour_manual(values=c("dark gray","red3"),guide=FALSE) +
  geom_hline(yintercept=obsticks,size=0.2) + 
  geom_vline(xintercept = predticks,size=0.2) + ggtitle('TF binding site: PAX8_B') +
  xlab('trans-acting gene') + ylab('target gene') + 
  coord_cartesian(xlim=c(predticks[1],predticks[23]),ylim=c(obsticks[1],obsticks[24]),expand=FALSE) +
  geom_text_repel(data=dplyr::filter(sorted2,txpath=='yes'),aes(label=paste(predname,":",obsname,sep=""))) + theme_classic(12) +
  scale_x_continuous(breaks=predchrticks,labels= c(1:16,"",18,"",20,"",22)) +
  scale_y_continuous(breaks=obschrticks,labels=c(1:16,"",18,"",20,"",22,"X"))
```


#SFig 1 top trans-acting pathways
```{r, fig.width=10,fig.height=5}
grid.arrange(sfig1a, sfig1b, nrow = 1)

png(filename=my.dir %&% 'SFig1_FUMA-trans.png',width = 600, height = 300)
grid.arrange(sfig1a, sfig1b, nrow = 1)
dev.off()

pdf(file=my.dir %&% 'SFig1_FUMA-trans.pdf',width=10,height=5)
grid.arrange(sfig1a, sfig1b, nrow = 1)
dev.off()

# tiff(filename=my.dir %&% 'SFig1_FUMA-trans.tiff', width = 6, height = 3, units = 'in', res = 300, compression = 'lzw')
# grid.arrange(sfig1a, sfig1b, nrow = 1)
# dev.off()
```

#SFig 2 top TF targets pathways
```{r, fig.width=10,fig.height=5}
grid.arrange(sfig2a, sfig2c, nrow = 1)

png(filename=my.dir %&% 'SFig2_FUMA-targets.png',width = 600, height = 300)
grid.arrange(sfig2a, sfig2c, nrow = 1)
dev.off()

pdf(file=my.dir %&% 'SFig2_FUMA-targets.pdf', width=10,height=5)
grid.arrange(sfig2a, sfig2c, nrow = 1)
dev.off()

# tiff(filename=my.dir %&% 'SFig2_FUMA-targets.tiff', width = 4, height = 4, units = 'in', res = 300, compression = 'lzw')
# grid.arrange(sfig2a, sfig2b, sfig2c, sfig2d, nrow = 2)
# dev.off()
```